---
title: "Donut and pie maps"
output: 
  bookdown::html_vignette2:
bibliography: '`r system.file("tmap.bib", package="tmap")`'
csl: "`r system.file('ieee.csl', package = 'tmap')`"
editor_options: 
  chunk_output_type: console
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  results = 'hide',
  fig.width = 8,
  comment = "#>"
)
hook_output <- knitr::knit_hooks$get("output")
knitr::knit_hooks$set(output = function(x, options) {
   lines <- options$output.lines
   if (is.null(lines)) {
     return(hook_output(x, options))  # pass to default hook
   }
   x <- unlist(strsplit(x, "\n"))
   more <- "..."
   if (length(lines)==1) {        # first n lines
     if (length(x) > lines) {
       # truncate the output, but add ....
       x <- c(head(x, lines), more)
     }
   } else {
     x <- c(more, x[lines], more)
   }
   # paste these lines together
   x <- paste(c(x, ""), collapse = "\n")
   hook_output(x, options)
 })

```


```{r, echo = FALSE, message = FALSE}
library(tmap)
library(tmap.glyphs)
tmap_options(scale = 0.75)
```

Glyphs in spatial data visualization are graphical symbols that represent data values at specific geographic locations. Each glyph can encode multiple data variables.

With the extension package [**tmap.glyphs**](https://github.com/r-tmap/tmap.glyphs/) (in development) glyphs can be created. Currently the donut, the pie and the flower glyphs are implemented. More will follow.

In this example we explore the proportion of low, middle and high incomes in Zuid-Holland, a province of the Netherlands (including The Hague and Rotterdam). We use the dataset `NLD_muni` from `tmap`, where we apply a filter and subsequently derive the percentage of middle income:

```{r, fig.height=6}
ZH_muni = NLD_muni[NLD_muni$province == "Zuid-Holland", ]
ZH_muni$income_middle = 100 - ZH_muni$income_high - ZH_muni$income_low
```


## Pie maps

Pie maps are (just) a special case of donut maps. 

```{r, fig.height=6}
tm_shape(ZH_muni) +
  tm_polygons() +
  tm_pies(parts = tm_vars(c("income_low", "income_middle", "income_high"), multivariate = TRUE),
    fill.scale = tm_scale_categorical(values = "-pu_gn_div"),			  
    size = "population",
    lwd = 1,
    size.scale = tm_scale_continuous(ticks = c(50000, 100000, 250000, 500000)))
```


Explanation:

* The visual variable `parts` is needed to determine the cut the pies. Note that a multivariate variable is required (see [vignette](https://r-tmap.github.io/tmap/articles/adv_multivariate)).
* The fill colors of the parts are specified via `fill.scale` (and not via `parts`). The used scale should be a categorical; each part represents a category.
* The visual variable `size` represents the overall sizes of the pies.


## Donut maps

Pies are just a specific case of the more general donuts.

```{r, fig.height=6}
tm_shape(ZH_muni) +
  tm_polygons() +
  tm_donuts(
  	parts = tm_vars(c("income_low", "income_middle", "income_high"), multivariate = TRUE),
    fill.scale = tm_scale_categorical(values = "-pu_gn_div"),			  
    size = "population",
    lwd = 1,
    size.scale = tm_scale_continuous(ticks = c(50000, 100000, 250000, 500000)))
```

Via the argument `options`, several properties of the donuts can be specified: 

```{r, fig.height=6}
tm_shape(ZH_muni) +
  tm_polygons() +
  tm_donuts(
  	parts = tm_vars(c("income_low", "income_middle", "income_high"), multivariate = TRUE),
    fill.scale = tm_scale_categorical(values = c("limegreen", "steelblue", "brown"), value.neutral = "grey50"),			  
    size = "population",
    lwd = 1,
    size.scale = tm_scale_continuous(ticks = c(50000, 100000, 250000, 500000)),
    options = opt_tm_donuts(start = 45, direction = -1, inner = 0.2, fill_hole = "orange"))
```

Explanation:

* `value.neutral` is used for legends of other visual variables (in this case size). See [vignette on neutral values](https://r-tmap.github.io/tmap/articles/adv_legends#neutral-values)
* For each tmap layer, `options` is used for option specific settings. The default list is obtained via `opt_tm_<layer>`.
